Lang-gen model simulations

### 2018-02-12 analysis/model-simulations.Rmd
add figure showing uniform prior simulation in detail

```{r}
save(sims.priors, sims.uncertain.thresholds, sims.fixed.thresholds,s1.simulations.relabeled, file = "../paper/cached_results/modelSims-priors_fixedT_uncertainT_speaker.RData")
```

>>>>

```{r saveResults}
#save(sims.priors, sims.uncertain.thresholds, sims.fixed.thresholds,s1.simulations.relabeled, file = "../paper/cached_results/modelSims-priors_fixedT_uncertainT_speaker.RData")
```

# Schematic (uniform prior) figure

```{r schematicModel}
schematicModel <- '
var round = function(x){
  return Math.round(x*100)/100
}

var isNegation = function(utt){
  return (utt.split("_")[0] == "not")
};

var avoidEnds = function(x){
  return x >= 1 ? 0.99 : x == 0 ? 0.01 : x
}

var lb = 0, ub = 1, diff = 0.05;
var bins = _.range(lb, ub + diff, diff)

var DiscreteBeta = function(a, b){
  Infer({model: function(){
    categorical({
      vs:bins,
      ps:map(function(x){
        Math.exp(Beta({a, b}).score(avoidEnds(x)))
      }, bins)
    })
  }})
}

var utterances = [
  "gen",
  "null"
];

var utterancePrior = Infer({model: function(){
  return uniformDraw(utterances)
}});

var meaning = function(words, state, threshold){
  return words == "gen" ? state > threshold :
  true
};

var speakerOptimality = 3;
var speakerOptimality2 = 1;

var fixedThetalistener0 = cache(function(utterance, threshold) {
  Infer({model: function(){
    var state = sample(DiscreteBeta(1, 1));
    var m = meaning(utterance, state, threshold);
    condition(m);
    return state;
  }})
}, 10000);

var greaterThanThresholdBins = _.range(lb, ub, diff)
var lessThanThresholdBins = _.range(lb+diff, ub+diff, diff)

var litInterpreter = cache(function(utterance) {
  Infer({model: function(){
    var threshold = uniformDraw(greaterThanThresholdBins)

    var state = sample(DiscreteBeta(1, 1));
    var m = meaning(utterance, state, threshold);
    condition(m);
    return state
  }})
}, 10000);

var speaker1 = cache(function(state) {
  Infer({model: function(){
    var utterance = sample(utterancePrior);
    var L0 = litInterpreter(utterance);
    factor(speakerOptimality*L0.score(state));
    return utterance;
  }})
}, 10000);
'

listenerCall <- '
_.fromPairs(map(function(u){
  var post = litInterpreter(u)
  return [u, post]
}, utterances))
'

speakerCall <- '
_.flatten(
    map(function(s){
      var speakProbs = speaker1(s)
       return {  
          state: s,
          "gen": Math.exp(speakProbs.score("gen")),
          "null": Math.exp(speakProbs.score("null"))
        }
    }, bins)
)
'

literalListenerCall <- '
_.flatten(_.flatten(
map(function(tH){
    var l0posterior = fixedThetalistener0("gen", tH)
    display(expectation(l0posterior))
    map(function(s){
       return {  
          state: s,
          theta:tH, 
          literalPosterior: Math.exp(l0posterior.score(s))
        }
    }, bins)
}, greaterThanThresholdBins)
))
'
```

### Fixed thresholds

```{r schematicFixed}
rs.wp.l0 <- webppl(paste(schematicModel, literalListenerCall,  sep = '\n'))

fig.l0.thresholds <- ggplot(rs.wp.l0, aes( x = state, 
                                           y = literalPosterior, 
                                       group = theta))+
  geom_line(size = 2)+
  geom_vline(aes(xintercept = theta), 
             color = 'darkred', size = 2)+
  #scale_color_solarized()+
  scale_x_continuous(breaks = c(0, 0.5, 1))+
  scale_y_continuous(breaks = c(0, 0.5, 1))+
  #scale_color_continuous(limits = c(0, 1), breaks = c(0, 1))+
  ggtitle("truth-functional threshold = ")+
  ylab("Literal listener posterior probability")+
  xlab("Degree of happiness")+
  facet_wrap(~theta)+
  theme(text = element_text(size = 16))
```

Integrate out threshold (full interpretation model)

```{r schematicUncertain}
rs.listener.wp <- webppl(paste(schematicModel, listenerCall,  sep = '\n'))

rs.listener.wp.tidy <- bind_rows(
  data.frame(rs.listener.wp$gen) %>% 
    mutate(utterance = "gen"),
  data.frame(rs.listener.wp$null) %>% 
    mutate(utterance = "null")
)

rs.listener.wp.tidy.samples <- get_samples(
  rs.listener.wp.tidy %>% rename(prob = probs), 10000)

ggplot(rs.listener.wp.tidy.samples, 
       aes( x = support,fill = utterance, color = utterance))+
  geom_density(alpha = 0.4, size = 1.3)+
  scale_fill_solarized()+
  scale_color_solarized()+
  xlab("prevalence")+
  ylab("Posterior probability density")+
  scale_x_continuous(breaks =c(0, 1))+
  scale_y_continuous(breaks = c(0, 2))+
  #guides(fill = F, color = F)+
  ggtitle("prevalence posterior")

#ggsave("figs/lassiterL1_posteriors_wCost1,5_alpha1.png", width = 6, height = 4)
```

```{r schematicEndorsement}
rs.wp <- webppl(paste(schematicModel, speakerCall,  sep = '\n'))

rs.tidy <- data.frame(rs.wp) %>%
  gather(utt, prob, -state)

fig.thresholds <- ggplot(rs.tidy %>%
                           mutate(utt = factor(utt,
                                               levels=c("null","gen"))), 
                         aes( x = state, y = prob, fill = utt))+
  geom_col( 
           color = 'black')+
  #facet_wrap(~happy_theta)+
  scale_x_continuous(breaks = c(0, 0.5, 1))+
  scale_y_continuous(breaks = c(0, 0.5, 1))+
  ggtitle("truth-functional threshold = ")+
  ylab("Speaker probability of saying 'happy'")+
  xlab("Degree of happiness")+
  theme(text = element_text(size = 16))+
  scale_fill_solarized()

fig.thresholds
#ggsave(fig.thresholds, 
 #      file = "figs/lassiterS1_uttXstateXtheta_wCost.pdf", width = 7, height = 7)
```

```{r}
save(rs.wp.l0, rs.listener.wp.tidy.samples, rs.tidy,
     file = "../paper/cached_results/modelSims-uniform_fixedT_uncertainT_speaker.RData")
```
